---
title: "HW 2 for P8105 - hlz2108"
output: github_document
---

The code chunk below loads libraries.

```{r Library_Setup}
library(tidyverse)
library(readxl)
```

##Problem 1

Reading in Mr. Trash Wheel dataset.

```{r Trash_Wheel_df}
trash_wheel_df =
  read_excel(
    "./data/Trash_Wheel.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols ("A:N")) %>%
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  ) %>% 
view
```

Read precipitation data!

```{r Precipitation_df}
precip_2018 = 
 read_excel(
    "./data/Trash_Wheel.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
 ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)
  
precip_2017 = 
 read_excel(
    "./data/Trash_Wheel.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
 ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation

```{r Combine_Precip}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df =
  bind_rows(precip_2018,precip_2017)

left_join(precip_df, month_df, by = "month")
```

This data set contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, including some specific kinds of trash. There are a total of `r nrow(trash_wheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. 

##Problem 2

Reading in and cleaning NYC Transit Data.

```{r NYC_Transit_df}
NYC_transit_df = 
  read_csv(
    "./data/NYC_Transit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>% 
view
```

mutate(entry = recode(entry, "yes" = T, "no" = F))

This data set contains information related to each entrance and exit for each subway station in NYC. In our final dataset, I have only included the following variables: `r names(NYC_transit_df)` and removed several others, such as entrance and exit locations. In cleaning the data, the entry variable was converted into a logical variable. There are a total of `r nrow(NYC_transit_df)` rows and `r ncol(NYC_transit_df)` columns in our final dataset. The data is not very tidy in the final dataset as there are many NA variables.

There are a total of r distinct(NYC_transit_df, station_name) distinct stations.

`r sum(pull(NYC_transit_df,ada), na.rm=FALSE)` stations are ADA compliant.

The proportion of station entrances/exits without vending allow entrance are 
 
##Problem 3

Reading in FiveThirtyEight Data.

```{r FiveThirtyEight_df}
pols_month_df = read_csv("./data/fivethirtyeight_datasets/pols-month.csv")
unemployment_df = read_csv("./data/fivethirtyeight_datasets/unemployment.csv")
snp_df = read_csv("./data/fivethirtyeight_datasets/snp.csv")
```

